\documentclass[a4paper]{tufte-handout}

\title{2. Qualitative Molecular Orbital Theory \thanks{Wayne~W.~Z. Yeo}}

\author[ESB]{Electronic States and Bonding\thanks{Course Instructors: Michael Bearpark, Joāo Pedro Malhaldo, Andreas Kafizas}}

%\date{28 March 2010} % without \date command, current date is supplied

%\geometry{showframe} % display margins for debugging page layout

\usepackage{graphicx} % allow embedded images
  \setkeys{Gin}{width=\linewidth,totalheight=\textheight,keepaspectratio}
  \graphicspath{{graphics/}} % set of paths to search for images
\usepackage{amsmath,amsthm,mathrsfs}  % extended mathematics
\usepackage[version=4]{mhchem}        % extended chemistry
\usepackage{xcolor}   % colors (like highlighting)
\usepackage{booktabs} % book-quality tables
\usepackage{units}    % non-stacked fractions and better unit spacing
\usepackage{multicol} % multiple column layout facilities
\usepackage{lipsum}   % filler text
\usepackage{fancyvrb} % extended verbatim environments
  \fvset{fontsize=\normalsize}% default font size for fancy-verbatim environments

% Standardize command font styles and environments
\newcommand{\doccmd}[1]{\texttt{\textbackslash#1}}% command name -- adds backslash automatically
\newcommand{\docopt}[1]{\ensuremath{\langle}\textrm{\textit{#1}}\ensuremath{\rangle}}% optional command argument
\newcommand{\docarg}[1]{\textrm{\textit{#1}}}% (required) command argument
\newcommand{\docenv}[1]{\textsf{#1}}% environment name
\newcommand{\docpkg}[1]{\texttt{#1}}% package name
\newcommand{\doccls}[1]{\texttt{#1}}% document class name
\newcommand{\docclsopt}[1]{\texttt{#1}}% document class option name

\setlength\fboxsep{7px}

\newenvironment{docspec}{\begin{quote}\noindent}{\end{quote}}% command specification environment

\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}
\newenvironment{justification} {\begin{proof}[Justification]} {\end{proof}}

\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{example}{Example}

\begin{document}

\maketitle% this prints the handout title, author, and date

\begin{abstract}
\noindent
We have very quickly exhausted the quantum chemical problems we can solve exactly.
Yet in a way, we have hardly done any chemistry. As we will see, we cannot solve the 
Schrödinger equation exactly for molecules, and we will need to resort to approximations.
With them, we can numerically solve the many-electron Schrödinger
equation for molecules, find the energies of their electronic states, and begin to
understand their bonding. \textbf{With these, we can make chemistry a computational science.}
\end{abstract}

%\printclassoptions

\newthought{Even the simplest molecule}, \ce{H2+} consists of three particles, and its
Schrödinger equation cannot be solved analytically. This simple molecule
will allows us to work out the basic ideas of what will become \textbf{molecular orbital theory}.

\section{Symmetry Elements}

\subsection{Reading Character Tables}
Each point group has its own \textbf{character table}.

\begin{example}[Finding irreducible representations of the central atom]
  Here, we find the symmetry labels for the \ce{p} and \ce{d} orbitals of \ce{Pt},
  the central atom in \ce{[PtCl4]^{2-}}.

  \begin{enumerate}
    \item First, you have to know the geometry of your compound. For example, \ce{[PtCl4]^{2-}} is square planar.
    \item The next step is to determine the point group of the compound from this geometry by identifying the symmetry elements that it posesses. \textit{Use a point group flow chart}. The point group of the \ce{[PtCl4]^{2-}} complex is $D_{4\mathrm{h}}$.
    \item Now, consult the character table of the point group we determined in the previous step.
    The character table of the $D_{4\mathrm{h}}$ point group is shown here:
  \end{enumerate}

  $$\begin{array}{c|cccccccccc|cc} \hline
  D_\mathrm{4h} & E & 2C_4 & C_2 & 2C_2' & 2C_2'' & i & 2S_4 & \sigma_\mathrm{h} & 2\sigma_\mathrm{v} & 2\sigma_\mathrm{d} & & \\ \hline
  \mathrm{A_{1g}} & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & & x^2+y^2,z^2 \\
  \mathrm{A_{2g}} & 1 & 1 & 1 & -1 & -1 & 1 & 1 & 1 & -1 & -1 & R_z & \\
  \mathrm{B_{1g}} & 1 & -1 & 1 & 1 & -1 & 1 & -1 & 1 & 1 & -1 & & x^2-y^2 \\
  
  \mathrm{B_{2g}} & 1 & -1 & 1 & -1 & 1 & 1 & -1 & 1 & -1 & 1 & & xy \\
  \mathrm{E_g} & 2 & 0 & -2 & 0 & 0 & 2 & 0 & -2 & 0 & 0 & (R_x,R_y) & (xz,yz) \\
  \mathrm{A_{1u}} & 1 & 1 & 1 & 1 & 1 & -1 & -1 & -1 & -1 & -1 & & \\
  \mathrm{A_{2u}} & 1 & 1 & 1 & -1 & -1 & -1 & -1 & -1 & 1 & 1 & z & \\
  \mathrm{B_{1u}} & 1 & -1 & 1 & 1 & -1 & -1 & 1 & -1 & -1 & 1 & & \\
  \mathrm{B_{2u}} & 1 & -1 & 1 & -1 & 1 & -1 & 1 & -1 & 1 & -1 & & \\
  \mathrm{E_u} & 2 & 0 & -2 & 0 & 0 & -2 & 0 & 2 & 0 & 0 & (x,y) & \\ \hline
  \end{array}$$

  Since the three p-orbitals point directly along the $x$, $y$, and $z$ axes, they show the same transformational properties as the basis vectors,
  and consequently get the same symmetry labels, i.e.

  \begin{align*}
  \left.\begin{array}{c}
  \mathrm{p}_x \\ \mathrm{p}_y \end{array}\right\} &\rightarrow \mathrm{E_u} \\
  \mathrm{p}_z &\rightarrow \mathrm{A_{2u}} \ .
  \end{align*}
  
  The five d-orbitals show the same transformational properties as their quadratic function counterparts and thus they get the symmetry labels, i.e.
  
  \begin{align*}
  \mathrm{d}_{xy} &\rightarrow \mathrm{B_{2g}} \\
  \left.\begin{array}{c}
  \mathrm{d}_{xz} \\ \mathrm{d}_{yz} \end{array}\right\} &\rightarrow \mathrm{E_g} \\
  \mathrm{d}_{z^2} &\rightarrow \mathrm{A_{1g}} \\
  \mathrm{d}_{x^2 - y^2} &\rightarrow \mathrm{B_{1g}} \ .
  \end{align*}
  
\end{example}

\subsection{Reading Mulliken Labels}

\begin{tabular}{ l l l }
  \hline
    \ce{A} & &singly degenerate, symmetric about $C_n$ axis ($+1$ in tables)  \\

    \ce{B} & &singly degenerate, antisymmetric about $C_n$ axis ($-1$ in tables)  \\

    \ce{E} & &doubly generate  \\

    \ce{T} & &triply generate  \\
    subscript 1 & e.g. \ce{A1} &symmetric about $C_2$ axis $\perp$ to $C_n$ axis or $\sigma_v$ if no $C_2$ present \\
    subscript 2 & e.g. \ce{A2} &antisymmetric about $C_2$ axis $\perp$ to $C_n$ axis or $\sigma_v$ if no $C_2$ present \\
    g & e.g. \ce{E_{2}}$_{\mathrm{g}}$ &\textit{gerade}, symmetric to inversion $i$ \\
    u & e.g. \ce{B_{2}}$_{\mathrm{u}}$ &\textit{ungerade}, antisymmetric to inversion $i$ \\
    superscript ' & e.g. \ce{A}' &symmetric to $\sigma_h$ \\
    superscript '' & e.g. \ce{E}'' &antisymmetric to $\sigma_h$ \\
  \hline
  \end{tabular}

\section{Symmetry Adapted Linear Combinations}

\subsection{Reducible Representations}
When we construct \textbf{symmetry adapted linear combinations} to build an MO diagram, we choose the $z$
axis as the \textit{principal axis}
\sidenote{Almost all character tables follow the conventions stated in Mulliken's paper "Report on Notation for the Spectra of Polyatomic Molecules" \textit{The Journal of Chemical Physics} 23, \textbf{1997} (1955).}
 (the axis of highest symmetry). 

\section{Appendix A: Symmetry Transformations}

\begin{definition}[Symmetry Transformation]
  A symmetry transformation is a change in our point of view that does not change the results of possible experiments.
  If an observer $O$ sees a system in a state represented by a ray $\mathscr{R}$ or $\mathscr{R}_{1}$ or $\mathscr{R}_{2} \ldots,$ then an equivalent observer $O^{\prime}$ who looks at the same system will observe it in a different state, represented by a ray $\mathscr{R}^{\prime}$ or $\mathscr{R}_{1}^{\prime}$ or $\mathscr{R}_{2}^{\prime} \ldots,$ respectively, but the two observers must find the same probabilities
$$
P\left(\mathscr{R} \rightarrow \mathscr{R}_{n}\right)=P\left(\mathscr{R}^{\prime} \rightarrow \mathscr{H}_{n}^{\prime}\right)
$$
\end{definition}




\bibliography{quantum}
\bibliographystyle{plainnat}



\end{document}
